---
title: "MATH 216 Data Science Final Project (Fall 2016)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(forcats)
library(stringr)
library(ggplot2)
library(forcats)
library(pander)
library(knitr)
library(ggthemes)
library(gridExtra)

setwd("~/Documents/Fall 2016/Data Sc/Final_Project")

# Load marriage preferences dataset
preferences_raw <- read_csv("Marriage_Preferences.csv") %>% 
  tibble::rownames_to_column(var="id")
```

---
output: 
  html_document: 
    keep_md: yes
---


* Name: Trisha Singh
* Project title: Caste-Based Mate Selection in Modern India
* Write-up on RPubs: <http://rpubs.com/tsingh/final_project>

**Abstract**: I examine how the prevalence of the caste system in India varies with socio-economic status. Specifically, I answer the question: “When seeking a marriage partner, is the importance given to the partner’s caste related to a person’s income or education?”

I find strong statistically significant relationships between socioeconomic status and attitude towards the caste system. Every $1000 increase in annual income is correlated with a 0.3 percent decrease in probability of having a preference for partner's caste. Additionally, I find that a person with a Master’s degree is 7.5 percentage points less likely and a person with a Professional degree is  15 percentage points less likely to have caste preferences than someone with a Bachelor’s degree. A person working in the public sector is 9 percentage points more likely to give importance to parnter's caste than a person working in the private sector.


```{r, echo=FALSE, message=FALSE, fig.width=12, fig.height=8}

preferences <- preferences_raw %>% 
  
  # Creating caste and sub caste binary variables
  mutate(partner_caste = ifelse(Caste == "\n Any Caste \n" |
                                  grepl("Caste no bar", Caste)==TRUE |
                                  Caste == "\n None \n" |
                                  Caste == "n/a", 0, 1)) %>% 
  mutate(partner_subcaste = ifelse(`Sub Caste` == "\n Any Sub Caste \n" |
                                     `Sub Caste` == "\n None \n" |
                                     `Sub Caste` == "n/a", 0, 1)) %>% 
  
  # Remove weird case (1 obs)
  filter(!(partner_caste==0 & partner_subcaste==1)) %>% 
  
  # must be hindu for caste/ subcaste to apply
  filter(Religion=="\n Hindu") %>%  
  
  # Cleaning income variable: only keeping income denoted in Rs
  filter(grepl("Rs", `Annual Income`)==TRUE) %>%
  mutate(mult_factor = ifelse(grepl("lakh", `Annual Income`)==TRUE, 100000, 
                              ifelse(grepl("crore", `Annual Income`)== TRUE, 10000000, 1))) %>% 
  mutate(income = str_sub(`Annual Income`, 8, 1000000)) %>% 
  mutate(income = str_replace_all(income, " lakhs", "")) %>% 
  mutate(income = str_replace_all(income, " crores", "")) %>% 
  mutate(income = str_replace_all(income, ",", "")) %>% 
  mutate(income = as.numeric(income)*mult_factor) %>% 
  mutate(income_dollar = income*0.015) %>% 
  
  # Cleaning education variable: Bachelor's, Masters, Professional, PhD
  mutate(education = fct_recode(Education, 
                                "Bachelor's" = "\n B.A.",
                                "Professional" = "\n Aeronautical Engineering",
                                "Professional" = "\n Aviation Degree",
                                "Bachelor's" = "\n B.A.M.S.",
                                "Bachelor's" = "\n B.Arch",
                                "Bachelor's" = "\n B.Com.",
                                "Bachelor's" = "\n B.Ed.",
                                "Bachelor's" = "\n B.L.",
                                "Bachelor's" = "\n B.M.M.",
                                "Bachelor's" = "\n B.Sc IT/ Computer Science",
                                "Bachelor's" = "\n B.Sc.",
                                "Bachelor's" = "\n B.Sc. Nursing",
                                "Bachelor's" = "\n B.Tech.",
                                "Bachelor's" = "\n BBA",
                                "Bachelor's" = "\n BCA",
                                "Bachelor's" = "\n BDS",
                                "Bachelor's" = "\n BE",
                                "Bachelor's" = "\n BFA",
                                "Bachelor's" = "\n BFM (Financial Management)",
                                "Bachelor's" = "\n BHM (Hotel Management)",
                                "Bachelor's" = "\n BHMS",
                                "Bachelor's" = "\n BPT",
                                "Bachelor's" = "\n BPharm",
                                "Bachelor's" = "\n Bachelor Degree in Arts / Science / Commerce",
                                "Bachelor's" = "\n Bachelor Degree in Engineering / Computers",
                                "Bachelor's" = "\n Bachelor Degree in Management",
                                "Professional" = "\n CA",
                                "Professional" = "\n CFA (Chartered Financial Analyst)",
                                "Professional" = "\n CS",
                                "Professional" = "\n ICWA",
                                "Professional" = "\n LL.B.",
                                "Professional" = "\n LL.M.",
                                "Masters" = "\n M.A.",
                                "Masters" = "\n M.Arch.",
                                "Masters" = "\n M.Pharm",
                                "Masters" = "\n M.Phil.",
                                "Professional" = "\n M.S.(Engg.)",
                                "Masters" = "\n M.Sc.",
                                "Masters" = "\n M.Sc. IT / Computer Science",
                                "Masters" = "\n M.Tech.",
                                "Masters" = "\n MBA",
                                "Professional" = "\n MBBS",
                                "Masters" = "\n MCA",
                                "Masters" = "\n MCom",
                                "Masters" = "\n MD / MS (Medical)",
                                "Masters" = "\n MDS",
                                "Masters" = "\n ME",
                                "Masters" = "\n MFM (Financial Management)",
                                "Masters" = "\n MHM  (Hotel Management)",
                                "Masters" = "\n MHRM (Human Resource Management)",
                                "Masters" = "\n MPT",
                                "Masters" = "\n Master Degree in Arts / Science / Commerce",
                                "Masters" = "\n Master Degree in Management",
                                "Masters" = "\n Master Degree in Medicine",
                                "Masters" = "\n Masters Degree in Engineering / Computers",
                                "Doctorate" = "\n PGDCA",
                                "Doctorate" = "\n PGDM",
                                "Doctorate" = "\n Ph.D."
                                )) %>% 
  # cleaning height
  mutate(height=as.numeric(str_sub(Height, str_length(Height)-5, str_length(Height)-2))) %>% 
  mutate(partner_height=as.numeric(str_sub(Height_1, str_length(Height_1)-5, str_length(Height_1)-2))) %>% 

# Clean parent contact
  mutate(parent_contact = ifelse(`Parent Contact`=="\n\t\t\t\t\t\tAvailable\n\t\t\t\t\t\t", 1, 0)) %>% 

# Relevel education
  mutate(education=fct_relevel(education, "Bachelor's")) %>% 

# Group job and relevel
  mutate(job = fct_recode(`Employed in`,
                          "Public Sector" = "\n Government/PSU",
                          "Private Sector" = "\n Private Sector",
                          "Other" = "\n Self Employed",
                          "Other" = "\n Business",
                          "Other" = "\n Defence")) %>% 
  mutate(job=fct_relevel(job, "Private Sector")) %>% 

# Clean age
  mutate(age = as.numeric(str_match_all(Age, "[0-9]+")))

# Baseline partner caste prefs

baseline_caste <- as.numeric(table(preferences$partner_caste)[2]/
  (table(preferences$partner_caste)[1]+table(preferences$partner_caste)[2]))

## INCOME DECILES ##

# Create income deciles
preferences$income_deciles <- preferences$income_dollar %>% 
  cut(.,
      breaks=quantile(., probs=seq(0,1, by=0.10), na.rm=TRUE), 
      include.lowest=TRUE) %>% 
  as.numeric()

preferences %>% 
  group_by(income_deciles) %>% 
  summarize(prop_caste=mean(partner_caste)) %>% 
  ggplot(aes(x=income_deciles, y=prop_caste)) + 
  geom_point() +
  geom_smooth(method="lm") +
  theme(text = element_text(family="Verdana"), 
             axis.text.x = element_text(angle=45, hjust = 1, face = "bold"), 
             axis.text.y = element_text(face = "bold"), 
             plot.title = element_text(size=20)) +
  labs(title = "Relationship between Income and Caste Preferences",
         y = "Proportion of people who stated caste prefs", x = "Income Deciles")
```

